---
title: "Exploring Hierarchical orthologous groups with roma"
author: "Klara Kaleb"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring Hierarchical orthologous groups with roma}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This little vignette shows you how to get started with the data available for HOGs in the `roma` package.

## The HOGs

Hierarchical orthologous groups (also known as HOGs) are sets of genes that are defined with respect to particular taxonomic ranges of interest[1]. They group genes that have descended from a single common ancestral genes in that taxonomic range. 

HOGs hold a lot of useful information and have many applications in various contexts, including inference of gene function, study of gene evolution dynamics and comparative genomics. Each HOG has a taxonomic range - within it, a given HOG can branch into constructs known as subHOGs which arise in an event of gene duplication.

HOGs can be retrived either by their hog id or by one of their members. Let's say we are interested in a gene that goes by the name of HUMAN22168 - below is the way to access its HOG.

```{r, warning=FALSE, message=FALSE}
library(roma)

hog <- getHOG(id="HUMAN22168")

getObjectAttributes(hog)

hog_id = getAttribute(hog,'hog_id')

parent_hogs = getAttribute(hog,'parent_hogs')

dim(parent_hogs)

parent_hog_id = parent_hogs$hog_id[[1]]

children_hogs = getAttribute(hog,'children_hogs')

child_hog_id_1  = children_hogs$hog_id[[1]]
child_hog_id_1

child_hog_id_2  = children_hogs$hog_id[[2]]
child_hog_id_2

child_hog_1 = getHOG(id=child_hog_id_1)

getAttribute(child_hog_1,"level")


```
The above constructs an object holding information for the information available for a HOG in the current OMA database. We can recognise that  HOG:0261495.1a is a subhog of its parent hog HOG:0261495 and that it further splits into 2 children subhogs, HOG:0261495.1a.1a and HOG:0261495.1a.1b respectively. We can also see that this split occured at the Neognathae level. 

We have just detected a gene duplication - it would be interesting to see if there is any gene differentiation as a consenquence. We can check this by looking at the member protein annotations for each subhog_id and perform a GO enrichment analysis on this by utilising the Bioconductor package topGO.

```{r, warning=FALSE, message=FALSE}
library(topGO)

child_hog_1_members = getHOG(id = child_hog_id_1, members=TRUE)
child_hog_2_members = getHOG(id = child_hog_id_2, members=TRUE)

# We can now directly use the above data to construct a topGO object for further analysis (such as GO enrichment)

annotations_1 = getOntologies(child_hog_1_members$members)
annotations_2 = getOntologies(child_hog_2_members$members)

annotations = c(annotations_1,annotations_2)

topGO = getTopGO(annotations, myInterestingGenes = names(annotations_1), format = "geneID2GO")


```

(1) Altenhoff, A., Gil, M., Gonnet, G., & Dessimoz, C. (2013). Inferring Hierarchical Orthologous Groups from Orthologous Gene Pairs PLoS ONE, 8 DOI: 10.1371/journal.pone.0053786