---
title: "Sequence Analysis with OmaDB"
author: "Klara Kaleb"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sequence Mapping with OmaDB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Another useful function of the OmaDB package is its functionality to exactly and partially match sequences. We will explore this further in this vignette. 

Let's say we have a sequence of interest that we want to map, in this case:

 sequence = "MNDPSLLGYPNVGPQQQQQQQQQQHAGLLGKGTPNALQQQLHMNQLTGIPPPGLMNNSDVHTSSNNNSRQLLDQLANGNANMLNMNMDNNNNNNNNNNNNNNNGGGSGVMMNASTAAVNSIGMVPTVGTPVNINVNASNPLLHPHLDDPSLLNNPIWKLQLHLAAVSAQSLGQPNIYARQNAMKKYLATQQAQQAQQQAQQQAQQQVPGPFGPGPQAAPPALQPTDFQQSHIAEASKSLVDCTKQALMEMADTLTDSKTAKKQQPTGDSTPSGTATNSAVSTPLTPKIELFANGKDEANQALLQHKKLSQYSIDEDDDIENRMVMPKDSKYDDQLWHALDLSNLQIFNISANIFKYDFLTRLYLNGNSLTELPAEIKNLSNLRVLDLSHNRLTSLPAELGSCFQLKYFYFFDNMVTTLPWEFGNLCNLQFLGVEGNPLEKQFLKILTEKSVTGLIFYLRDNRPEIPLPHER"

We can pass it to the mapSequence() function which returns a list of targets. From this list we can then construct protein objects for which we can obtain further infromation - such as its oma group or its domains. The example response object, generated via mapSequence(sequence), is below.

```{r, warning=FALSE, message=FALSE}
library(OmaDB)

load('../data/sequence_map.rda')

getObjectAttributes(sequence_map)

targets = getAttribute(sequence_map,'targets')

length(targets) 

protein = targets[[1]][['entry_url']]

```

One can also directly obtain GO annotations for a given query sequence (protein$ogene_ontology), which results in an object as below:

```{r, warning=FALSE, message=FALSE}

load('../data/sequence_annotation.rda')

sequence_annotation 

```

